Wavefield Reconstruction Inversion: an example

### 2020-02-14 repro/papers/wri1d/paper.tex
first draft

\begin{abstract}
  Full waveform inversion, or nonlinear least squares data-fitting
  driven by wave equation simulation, has proven its ability to
  extract detailed maps of earth structture from near-surface seismic
  observations. However, the least squares error function at the heart
  of this method tends to develop a high degree of nonconvexity, so
  that local optimization methods (the only numerical methods feasible
  for field-scale problems) tend to fail, unless provided with initial
  estimates of a quality not always available. A number of alternative
  optimization principles have been advanced that promise some degree
  of release from the multimodality of Full Waveform Inversion,
  amongst them Wavefield Reconstruction Inversion, the focus of this
  paper. Applied to a simple 1D acoustic transmission problem, both
  Full Waveform and Wavefield Reconstruction Inversion methods reduce
  to minimization of explicitly computable functions, in an asymptotic
  sense. The analysis presented here shows that Wavefield
  Reconstruction Inversion can be vulnerable to the same
  ``cycle-skipping'' failure mode as Full Waveform Inversion.
\end{abstract}

@@

The objective of FWI is to find $c$ and $f$, given $d$, so that
$S[c]f \approx d$. A typical method for enforcing this requirement is
the minimization of an objective misfit measure (objective, for
short), the most common choice being the square norm of a Hilbert
space in which the data is presumed to reside:
\begin{equation}
  \label{eqn:deffwi}
  \mbox{Given } d, \mbox{ find }c \mbox{ and }f \mbox{ to
    minimize }
  \|d -  S[c]f\|^2
\end{equation}
in which $\|\cdot\|$ is the norm in a suitable Hilbert space. This
approach was first suggested in the 1980's
(\cite{BamChavLai:79,Tara:84a,KolbColLai:86,Crasetal:90}, and many
other papers since then). Usually some form of regularization is
applied, to compensate for poorly determined aspects of $c$ and/or
$f$, as is explained in Tarantola's influential book
\cite[]{Tarantola:05}. Also, $f$ may be constrained in one way or
another to embody characteristics of field energy sources, or even
regarded as known (an example of this is given below).

@@

This topic of this paper is one of these alternative approaches,
Wavefield Reconstruction Inversion (``WRI''). WRI was introduced by \cite{LeeuwenHerrmannWRI:13}, and further
developed by \cite{LeeuwenHerrmann:16,WangYingst:SEG16} and other
authors. It is based on the presumption that the correct source (right
hand side in the equations of motion) $q$ is
known, and combines a penalty for data misfit with a penalty for
failing to solve the equations of motion with the correct right hand side:
\begin{equation}
  \label{eqn:defwri}
  \mbox{Given } d \mbox{ and }q, \mbox{ find }c \mbox{ and }f \mbox{ to
    minimize }
  \|d -  S[c]f\|^2+\alpha^2\|f-q\|^2.
\end{equation}
The penalty parameter $\alpha$ may be chosen small, and examples given
by \cite{LeeuwenHerrmannWRI:13} and others suggest that the problem
defined in display \ref{eqn:defwri} is less likely to develop
uninformative local minima if $\alpha$ is small. This might be so for
several reasons: one is that when less weight is
put on making the residual in the wave equation ($f-q$) small, 
$f$ may be chosen to make the data residual $d-S[c]f$ small
instead. One aspect of failure to predict the arrival times of waves
accurately is that small data residual is then difficult to
achieve. By maintaining fidelity to
the data, local minimization of the WRI objective might provide a
spurious-minimum-free path to a satisfactory estimate of $c$.

@@

The formulation of the inverse problem via least-squares requires a choice of Hilbert space structure for the domain and range of $S[c]$. A natural choice is
\begin{equation}
  \label{eqn:defmoddom}
  S[c]: L^2([z_{\rm min},z_{\rm max}] \times \bR) \rightarrow L^2[0,T]
\end{equation}
As it will turn out, the choice of spatial interval $[z_{\rm min},z_{\rm max}]$ is arbitrary, so long as it has positive length and contains $z_s$ in its interior.

For homogeneous ($z$-independent) $\kappa$, Appendix A provides an explicit expression:
\begin{equation}
  \label{eqn:defmodhom}
  S[c]f(t) = \frac{1}{2c}\int_{z_{\rm min}}^{z_{\rm max}} dz f\left(z,t - \frac{|z_r-z|}{c}\right) 
\end{equation}
From this expression it is simple to see that $S[c]$ is a bounded
operator with the domain and range described in
\ref{eqn:defmoddom}. The role of compact support in $z$ in ensuring
boundedness is also evident. % It is also evident from that limiting source to have bounded support in $z$ is necessary make $S[c]$ bounded.

With this framework, the data is presumed to lie in the range space of $S[c]$: $d \in L^2[0,T]$.

The version of FWI discussed here presumes that the source field corresponding to the data $d$ is supported at a point $z_s \ne z_r$, and is known. This field depends on a function of time (``wavelet'') $w \in L^2[0,T]$. Formally, the resulting acoustic field satisfies the system \ref{eqn:awe1d} with $f=w(t)\delta(z-z_s)$. Inserting this expression in the explicit expression \ref{eqn:defmodhom}, obtain
\begin{equation}
  \label{eqn:defmodpt}
  S[c](w\delta(\cdot-z_s)) = \frac{1}{2c}w\left(t - \frac{|z_r-z_s|}{c}\right) = S_p[c]w(t)
\end{equation}
Thus for this restricted class of source fields, the FWI objective can be redefined as
\begin{equation}
  \label{eqn:deffwipt}
  J_{\rm FWI}[c,d.w] = \frac{1}{2}\|S_p[c]w -d\|^2.
\end{equation}
$S_p[c]$ is a bounded operator with domain $L^2(\bR)$ and range $L^2([0,T])$.

@@

To understand the non-convexity phenomenon and the
relation of the various scales, it is advantageous to introduce a
family of source wavelets, depending on a parameter $\lambda$ playing
the role of wavelength:
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right).
\end{equation}
In the definition \ref{eqn:wfam}, the ``mother wavelet'' $w_1\in
C_0^{\infty}(0,T)$, and the scaling is chosen so fhat
$\|\wl\|_{L^2(\bR)}$ is independent of $\lambda>0$.

@@

Choose a target wave velocity $c_* \in [c_{\rm min},c_{\rm
  max}]$. Introduce a family of consisten data $\dl$, generated by
$c_*$ and the wavelet family $\wl$:
\begin{equation}
  \label{eqn:dfam}
  \dl = S_p[c_*]\wl
\end{equation}
and a corresponding family of FWI objectives:
\[
  J_{\rm FWI}[c,\dl,\wl] = \frac{1}{2}\|\dl-S_p[c]\wl\|^2
\]
\begin{equation}
  \label{eqn:fwiexpl}
  =
  \int_0^T\,dt\,\left|\frac{1}{2c_*}\wl\left(t-\frac{z_s-z_r}{c_*}\right)
    - \frac{1}{2c}\wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2
\end{equation}
Note that $\mbox{supp }\wl \subset [0,\lambda T]$, so
\begin{equation}
  \label{eqn:wlsupps}
  \mbox{supp }S_p[c]\wl \subset \left[\frac{z_s-z_r}{c},\lambda
    T+\frac{z_s-z_r}{c}\right] \cap [0,T]
\end{equation}
The transit time condition \ref{eqn:time} implies that there exists
$\lambda_*>0$ so that for $\lambda < \lambda_*$,
\[
  \lambda T+\frac{z_s-z_r}{c} < T
\]
for all admissible $c$. That is, 
for $\lambda <\lambda_*$
\begin{equation}
  \label{eqn:constnorm}
  \|S_p[c]\wl\|^2 = \frac{1}{4c^2}\int_{-\infty}^{\infty}\,dt\,\left|
    \wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2 =
  \frac{\|w_1\|^2}{4c^2}.
\end{equation}

Recall that the object of this study is the global behaviour of
objective functions for velocity estimation: in this context, that
means the behaviour for $c$ far from $c_*$.  Pick $\delta c >0$ for
which $[c_*-\delta c, c_* + \delta c] \subset [c_{\rm min},c_{\rm
  max}]$, that is, that the set of admissible $c$ further than $\delta
c$ from $c_*$ is non-void.

Then there is $\lambda_{**}$ depending on $\delta c, T, and
|z_s-z_r|$ so that if $|c-c_*|>\delta c$ and $\lambda<\lambda_{**}$, then
\[
  \left|\frac{|z_s-z_r|}{c}-\frac{|z_s-z_r|}{c_*}\right| > 2\lambda T,
\]
that is, the infima of the supports of $S_p[c]\wl, \dl=S_p[c_*]\wl$
are further apart than the lengths of these supports. In that case,
necessarily
\[
  \mbox{supp }S_p[c]\wl \cap \mbox{supp }S_p[c_*]\wl = \emptyset.
\]
Therefore $S_p[c]\wl$ and $S_p[c_*]\wl$ are orthogonal in $L^2[0,T]$,
and
\begin{equation}
  \label{eqn:fwiscrewed}
  J_{\rm FWI,\lambda}[c] =
  \frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2.
\end{equation}
Amongst other consequences, one immediately deduces from the
expression \ref{eqn:fwiscrewed} the non-convexity result:

\begin{theorem}
  \label{thm:thm1}
  There exists $C>0$, $\lambda_0$ so that for $\lambda <
  \lambda_0$, the minimizer of $J_{\rm FWI, \lambda}[c]$ on the
  complement of $[c_*-C\lambda, c_*+C\lambda]$ is $c=c_{\rm max}$.
\end{theorem}

That is, outside of a neighborhood of width proportional to a
wavelength, minimization of $J_{\rm FWI}$ yields a local minimizer far from
the target $c_*$ which generates the (noise-free) data.

@@

\cite{LeeuwenHerrmann:16} used the variable projection method
\cite[]{GolubPereyra:03}, eliminating the source $f$ in the inner step
and updating the bulk modulus (or an equivalent quantity) in the outer
step. That is, the problem \ref{eqn:defwri} is equivalent to
minimization of
\begin{equation}
\label{eqn:defvpm}
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_f \frac{1}{2}(\| d-S[c]f\|^2+\alpha^2\|f-q\|^2)
\end{equation}
over $c$. This is the approach that I shall pursue here.

@@

``Most'' source fields $f$ are non-radiating,
that is, $S[c]f=0$. The orthocomplement of the subspace of
non-radiating sources is the range of the adjoint operator
$S[c]^T$. This observation suggest that the sources may be re-parametrized as images of
$S[c]^T$. Write $f = g + q$: then the second term in
the objective displayed in \ref{eqn:defwri} is simply $\|g\|^2$.
Decompose $g = S[c]^Te + n$, in
which $e$ is the same type of object as $d$ and $S[c]n=0$ (that
is, $n$ is a non-radiating
source), and note that the decomposition is orthogonal. Then
\[
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_{e,n} \frac{1}{2}(\| d-S[c](S[c]^Te+q)\|^2+\alpha^2(\|S[c]^Te\|^2+\|n\|^2))
  \]
\begin{equation}
  \label{eqn:defvpmred}
   \mbox{min}_{e} \frac{1}{2}(\|r-S[c]S[c]^Te\|^2+\alpha^2\|S[c]^Te\|^2)
 \end{equation}
 in which $r[c]=d-S[c]q$ is the data residual.

This reformulation has some computational advantages \cite[]{WangYingst:SEG16,Herrmann:SEG19}, but also leads to a useful analytic transformation of the WRI problem. The minimizer on the RHS of equation \ref{eqn:defvpmred} is the solution $e=e_{\alpha}[c]$ of the normal equation
\[
 ( (S[c]S[c]^T)^2 + \alpha^2S[c](S[c]^T)e = S[c]S[c]^Tr[c]
\]
whence
\[
  S[c]S[c]^Te_{\alpha}[c] = S[c]S[c]^T(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\]
Assume that $S[c]S[c]^T$ is invertible - this is in fact for the problem outlined in the introduction, as will be verified shortly. Then
\begin{equation}
  \label{eqn:norsol}
  e_{\alpha}[c]=(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\end{equation}
Consequently
\[
  J_{\rm WRI}[c,d,q] =
  \frac{1}{2}(\|r[c]-S[c]S[c]^Te_{\alpha}[c]\|^2+\alpha^2\|S[c]^Te_{\alpha}[c]\|^2)
\]
\[
  = \frac{1}{2}\left(\|r[c]-S[c]S[c]^T
    (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2 \right.
\]
\[
  \left.+\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  S[c]S[c]^T (S[c]S[c]^T+\alpha^2I)^{-1}r[c]
  \rangle\right)
\]
\[
  =\frac{1}{2}\left(\|\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2\right.
\]
\[
  \left.
  +\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  -\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\rangle\right.
\]
\[
  + \left.\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle\right)
\]
\[
  =  \frac{\alpha^2}{2}
\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle 
\]
That is,
\begin{equation}
  \label{eqn:defwrialt}
  J_{\rm WRI}[c,d,q] = \frac{1}{2}\langle r[c], W_{\alpha}[c] r[c]\rangle
\end{equation}
with
\begin{equation}
  \label{eqn:defwriwt}
  W_{\alpha}[c] = \frac{\alpha^2}{2}(S[c]S[c]^T+\alpha^2I)^{-1}
\end{equation}
This remarkable identity shows that the WRI objective function is a {\em weighted norm of the data residual $r[c]$}. Note that this conclusion depends only on the presumption that the operator $S[c]S[c]^T$ is invertible (or, more precisely, injective), a condition that will be verified below for the problem considered in this paper, and which holds in many other settings in which WRI has been advocated.

@@

Define
\begin{equation}
  \label{eqn:defbeta}
\beta = \frac{z_{\rm max}-z_{\rm min}}{c_*^2} - 4 \alpha^2
\end{equation}
\begin{theorem}
  \label{thm:thm2}
  There exist $C>0$, $\lambda_0$ so that for $\lambda <
  \lambda_0$, the minimizer of $J_{\rm WRI}[c,\dl,\wl \delta(\cdot-z_s)]$ on the
  complement of $[c_*-C\lambda, c_*+C\lambda]$ is
  \begin{itemize}
  \item $c=c_{\rm max}$ if $\beta<0$;
  \item $c=c_{\rm min}$ if $\beta>0$;
  \item any $c < c_*-C\lambda$ or $>c_*+C\lambda$ if $\beta=0$.
  \end{itemize}
\end{theorem}
\begin{proof}
  From \ref{eqn:wriscrewed}, \ref{eqn:wriwthom}, and
  \ref{eqn:fwiscrewed},
  \[
    J_{\rm WRI}[c,\dl,\wl\delta(\cdot-z_s)] =
    \frac{\alpha^2}{2}\left(\frac{z_{\rm max}-z_{\rm
          min}}{4c^2}+\alpha^2\right)^{-1}\frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2
  \]
  \[
    =
    \frac{\alpha^2}{4}\frac{1+\frac{c^2}{c_*^2}}{z_{\rm 
        max}-z_{\rm min}+4c^2\alpha^2}
  \]
  This linear fractional function of $c^2$ is increasing, decreasing,
  or constant if $\beta>0, \beta<0$ or $\beta=0$, respectively.
\end{proof}

@@

\section{Discussion}

(TBA)

\section{Conclusion}

(TBA)

>>>>

\begin{abstract}
  Full waveform inversion, or nonlinear least squares data-fitting
  driven by wave equation simulation, has proven its ability to
  extract detailed maps of earth structure from near-surface seismic
  observations. However, the least squares error function at the heart
  of this method tends to develop a high degree of nonconvexity, so
  that local optimization methods (the only numerical methods feasible
  for field-scale problems) tend to fail, unless provided with initial
  estimates of a quality not always available. A number of alternative
  optimization principles have been advanced that promise some degree
  of release from the multimodality of Full Waveform Inversion,
  amongst them Wavefield Reconstruction Inversion, the focus of this
  paper. Applied to a simple 1D acoustic transmission problem, both
  Full Waveform and Wavefield Reconstruction Inversion methods reduce
  to minimization of explicitly computable functions, in an asymptotic
  sense. The analysis presented here shows that Wavefield
  Reconstruction Inversion can be vulnerable to the same
  ``cycle-skipping'' failure mode as Full Waveform Inversion.
\end{abstract}

@@

The objective of FWI is to find $c$ and $f$, given $d$, so that
$S[c]f \approx d$. A typical method for enforcing this requirement is
the minimization of an objective misfit measure (objective, for
short), the most common choice being the square norm of a Hilbert
space in which the data is presumed to reside:
\begin{equation}
  \label{eqn:deffwi}
  \mbox{Given } d, \mbox{ find }c \mbox{ and }f \mbox{ to
    minimize }
  \|d -  S[c]f\|^2
\end{equation}
in which $\|\cdot\|$ is the norm in a suitable Hilbert space. This
approach was first suggested in the 1980's
(\cite[]{BamChavLai:79,Tara:84a,KolbColLai:86,Crasetal:90}, and many
other papers since then). Usually some form of regularization is
applied, to compensate for poorly determined aspects of $c$ and/or
$f$, as is explained in Tarantola's influential book
\cite[]{Tarantola:05}. Also, $f$ may be constrained in one way or
another to embody characteristics of field energy sources, or even
regarded as known (an example of this is given below).

@@

This topic of this paper is one of these alternative approaches,
Wavefield Reconstruction Inversion (``WRI''). WRI was introduced by \cite{LeeuwenHerrmannWRI:13}, and further
developed by \cite{LeeuwenHerrmann:16,WangYingst:SEG16} and other
authors. It is based on the presumption that the correct source (right
hand side in the equations of motion) $q$ is
known, and combines a penalty for data misfit with a penalty for
failing to solve the equations of motion with the correct right hand side:
\begin{equation}
  \label{eqn:defwri}
  \mbox{Given } d \mbox{ and }q, \mbox{ find }c \mbox{ and }f \mbox{ to
    minimize }
  \|d -  S[c]f\|^2+\alpha^2\|f-q\|^2.
\end{equation}
Examples given
by \cite{LeeuwenHerrmannWRI:13} and others suggest that the problem
defined in display \ref{eqn:defwri} is less likely to develop
uninformative local minima if $\alpha$ is small. This might be so for
several reasons: one is that when less weight is
put on making the residual in the wave equation ($f-q$) small, 
$f$ may be chosen to make the data residual $d-S[c]f$ small
instead. One aspect of failure to predict the arrival times of waves
accurately is that small data residual is then difficult to
achieve. By maintaining fidelity to
the data, local minimization of the WRI objective might provide a
spurious-minimum-free path to a satisfactory estimate of $c$.

@@

The formulation of the inverse problem via least-squares requires a choice of Hilbert space structure for the domain and range of $S[c]$. A natural choice is
\begin{equation}
  \label{eqn:defmoddom}
  S[c]: L^2([z_{\rm min},z_{\rm max}] \times \bR) \rightarrow L^2[0,T]
\end{equation}
That is, the support of f will be assumed to lie in the strip $[z_{\rm min},z_{\rm max}] \times \bR$.
As it will turn out, the choice of spatial interval $[z_{\rm min},z_{\rm max}]$ is arbitrary, so long as it has positive length and contains $z_s$ in its interior.

For homogeneous ($z$-independent) $\kappa$, Appendix A provides an explicit expression:
\begin{equation}
  \label{eqn:defmodhom}
  S[c]f(t) = \frac{1}{2c}\int_{z_{\rm min}}^{z_{\rm max}} dz f\left(z,t - \frac{|z_r-z|}{c}\right) 
\end{equation}
From this expression it is simple to see that $S[c]$ is a bounded
operator with the domain and range described in
\ref{eqn:defmoddom}. The role of compact support in $z$ in ensuring
boundedness is also evident.
%It is also evident from that limiting source to have bounded support in $z$ is necessary make $S[c]$ bounded.

With this framework, the data is presumed to lie in the range space of
$S[c]$: $d \in L^2[0,T]$. It is simple to verify from the identity
\ref{eqn:defmodhom} that $S[c]$ is also surjective. That is, any data
at all can be fit by $S[c]$ with an appropriate choice of input. This
observation is important in the development of WRI, to be explained
below.

The version of FWI discussed here presumes that the source field
corresponding to the data $d$ is supported at a point $z_s \ne z_r$,
and is known. This field depends on a function of time (``wavelet'')
$w \in L^2(\bR)$. Formally, the resulting acoustic field satisfies the
system \ref{eqn:awe1d} with $f=w(t)\delta(z-z_s)$. Inserting this
expression in the explicit expression \ref{eqn:defmodhom}, obtain
\begin{equation}
  \label{eqn:defmodpt}
  S[c](w\delta(\cdot-z_s)) = \frac{1}{2c}w\left(t - \frac{|z_r-z_s|}{c}\right) = S_p[c]w(t)
\end{equation}
Thus for this restricted class of source fields, the FWI objective can be redefined as
\begin{equation}
  \label{eqn:deffwipt}
  J_{\rm FWI}[c,d.w] = \frac{1}{2}\|S_p[c]w -d\|^2.
\end{equation}
$S_p[c]$ is a bounded operator with domain $L^2(\bR)$ and range $L^2([0,T])$.

@@

To understand the non-convexity phenomenon and the
relation of the various scales, it is advantageous to introduce a
family of source wavelets, depending on a parameter $\lambda$, having
dimensions of time and playing
the role of wavelength:
\begin{equation}
  \label{eqn:wfam}
  \wl(t) = \frac{1}{\sqrt{\lambda}}w_1\left(\frac{t}{\lambda}\right).
\end{equation}
In the definition \ref{eqn:wfam}, the ``mother wavelet'' $w_1\in
C_0^{\infty}(0,1)$, and the scaling is chosen so fhat
$\|\wl\|_{L^2(\bR)}$ is independent of $\lambda>0$.

@@

Choose a target wave velocity $c_* \in [c_{\rm min},c_{\rm
  max}]$. Introduce a family of consisten data $\dl$, generated by
$c_*$ and the wavelet family $\wl$:
\begin{equation}
  \label{eqn:dfam}
  \dl = S_p[c_*]\wl
\end{equation}
and a corresponding family of FWI objectives:
\[
  J_{\rm FWI}[c,\dl,\wl] = \frac{1}{2}\|\dl-S_p[c]\wl\|^2
\]
\begin{equation}
  \label{eqn:fwiexpl}
  =
  \int_0^T\,dt\,\left|\frac{1}{2c_*}\wl\left(t-\frac{z_s-z_r}{c_*}\right)
    - \frac{1}{2c}\wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2
\end{equation}
Note that $\mbox{supp }\wl \subset [0,\lambda]$, so
\begin{equation}
  \label{eqn:wlsupps}
  \mbox{supp }S_p[c]\wl \subset \left[\frac{z_s-z_r}{c},\lambda
    +\frac{z_s-z_r}{c}\right] \cap [0,T]
\end{equation}
The transit time condition \ref{eqn:time} implies that there exists
$\lambda_0>0$ so that for $\lambda < \lambda_0$,
\[
  \lambda +\frac{z_s-z_r}{c} < T
\]
for all admissible $c$. That is, 
for $\lambda <\lambda_0$, $\mbox{supp }S_p[c]\wl \subset(0,T)\mbox{
  for }c \in [c_{\rm min},c_{\rm max}]$, and 
\begin{equation}
  \label{eqn:constnorm}
  \|S_p[c]\wl\|^2 = \frac{1}{4c^2}\int_{-\infty}^{\infty}\,dt\,\left|
    \wl\left(t-\frac{z_s-z_r}{c}\right)\right|^2 =
  \frac{\|w_1\|^2}{4c^2}.
\end{equation}

Recall that the object of this study is the global behaviour of
objective functions for velocity estimation: in this context, that
means the behaviour for $c$ far from $c_*$. Define
\begin{equation}
  \label{eqn:defL}
  L = \frac{2 c_{\rm max}^2}{|z_s-z_r|}.
\end{equation}
Then if $|c_*-c| > L\lambda$,
\[
  \left|\frac{|z_s-z_r|}{c}-\frac{|z_s-z_r|}{c_*}\right|
  =\frac{|c-c_*||z_s-z_r|}{cc_*} 
\]
\[
\ge \frac{|c-c_*||z_s-z_r|}{c_{\rm max}^2} > L\lambda
\frac{|z_s-z_r|}{c_{\rm max}^2} = 2\lambda.
\]
That is, the infima of the supports of $S_p[c]\wl, \dl=S_p[c_*]\wl$
are further apart than the lengths of these supports. In that case,
necessarily
\[
  \mbox{supp }S_p[c]\wl \cap \mbox{supp }S_p[c_*]\wl = \emptyset.
\]
Therefore $S_p[c]\wl$ and $S_p[c_*]\wl$ are orthogonal in $L^2[0,T]$,
and
\begin{equation}
  \label{eqn:fwiscrewed}
  J_{\rm FWI,\lambda}[c] =
  \frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2.
\end{equation}
Amongst other consequences, one immediately deduces from the
expression \ref{eqn:fwiscrewed} the non-convexity result:

\begin{theorem}
  \label{thm:thm1}
  For $L>0$ given by equation \ref{eqn:defL} and $\lambda <
  \lambda_0$, the minimizer of $J_{\rm FWI}[c,\dl,\wl]$ on the
  complement of $[c_*-L\lambda, c_*+L\lambda]$ is $c=c_{\rm max}$.
\end{theorem}

That is, outside of a neighborhood of width proportional to a
wavelength, minimization of $J_{\rm FWI}$ yields a local minimizer far from
the target velocity $c_*$ that generates the (noise-free) data.

@@

\cite{LeeuwenHerrmann:16} used the variable projection method
\cite[]{GolubPereyra:03}, eliminating the source $f$ in the inner step
and updating the bulk modulus (or an equivalent quantity) in the outer
step. That is, the problem \ref{eqn:defwri} is equivalent to
minimization of
\[
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_f \frac{1}{2}(\| d-S[c]f\|^2+\alpha^2\|f-q\|^2)
\]
over $c$. This is the approach that I shall pursue here.

It will turn out to be convenient to define the residual with the
target source $q$ as $r[c]=d-S[c]q$, and set $g=f-q$. Then this
definition can be rewritten as
\begin{equation}
\label{eqn:defvpm}
  J_{\rm WRI}[c,d,q] =\mbox{min}_g \frac{1}{2}(\| r[c]-S[c]g\|^2+\alpha^2\|g\|^2)
\end{equation}

@@

``Most'' source fields $f$ are non-radiating, that is, $S[c]f=0$, and
such sources contribute nothing to the data fit term in the definition
of $J_{\rm WRI}$. If the domain and range of $S$ were finite
dimensional (which of course they are, after discretization), then the
Fundamental Theorem of Linear Algebra identifies the null space of
$S[c]$ (the non-radiating sources) as the orthocomplement of the range
of the transpose $S[c]^T$ \cite[]{Strang:93}. In the
infinite-dimensional setting of this paper, the Closed Range Theorem
\cite[]{Yosida} states that the same is true if the range of $S[c]$ is
closed. There are various ways to ensure this property, but the
simplest is relevant here:
\begin{quote}
  For all admissible models $c$, $S[c]$ is surjective.
\end{quote}
That is, {em any data can be fit exactly using the extended model
  space (the domain of $S$)}.
This property is a characteristic of of extended modeling
methods \cite[]{geoprosp:2008}: the ability to fit any data appears to
be a essential for such methods to produce objectives
without spurious local minima. It holds for the problems for which WRI
has been advocated. For the simple model problem considered
here, surjectivity follows from the explicit expression for $S[c]f$,
as was noted in the discussion following equation \ref{eqn:defmodhom}.

Assuming that $S[c]$ is surjective for any admissible $c$, the
orthocomplement of the subspace of non-radiating sources is the range
of the adjoint operator $S[c]^T$. In the
definition \ref{eqn:defvpm}, decompose $g = S[c]^Te + n$, in
which $e$ is the same type of object as $d$ and $S[c]n=0$ (that is,
$n$ is a non-radiating source), and note that the decomposition is
orthogonal. Then
\[
  J_{\rm WRI}[c,d,q] =
  \mbox{min}_{e,n} \frac{1}{2}(\| d-S[c](S[c]^Te+q)\|^2+\alpha^2(\|S[c]^Te\|^2+\|n\|^2))
  \]
\begin{equation}
  \label{eqn:defvpmred}
   \mbox{min}_{e} \frac{1}{2}(\|r[c]-S[c]S[c]^Te\|^2+\alpha^2\|S[c]^Te\|^2).
 \end{equation}

This reformulation has some computational advantages \cite[]{WangYingst:SEG16,Herrmann:SEG19}, but also leads to a useful analytic transformation of the WRI problem. The minimizer on the RHS of equation \ref{eqn:defvpmred} is the solution $e=e_{\alpha}[c]$ of the normal equation
\[
 ( (S[c]S[c]^T)^2 + \alpha^2S[c](S[c]^T)e = S[c]S[c]^Tr[c]
\]
whence
\[
  S[c]S[c]^Te_{\alpha}[c] = S[c]S[c]^T(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\]
Since the null space of $S[c]$ is orthogonal to the range of $S[c]^T$
under the surjectivity assumption, $S[c]S[c]^T$ is injective, whence
\begin{equation}
  \label{eqn:norsol}
  e_{\alpha}[c]=(S[c]S[c]^T+\alpha^2I)^{-1}r[c]
\end{equation}
Consequently
\[
  J_{\rm WRI}[c,d,q] =
  \frac{1}{2}(\|r[c]-S[c]S[c]^Te_{\alpha}[c]\|^2+\alpha^2\|S[c]^Te_{\alpha}[c]\|^2)
\]
\[
  = \frac{1}{2}\left(\|r[c]-S[c]S[c]^T
    (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2 \right.
\]
\[
  \left.+\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  S[c]S[c]^T (S[c]S[c]^T+\alpha^2I)^{-1}r[c]
  \rangle\right)
\]
\[
  =\frac{1}{2}\left(\|\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\|^2\right.
\]
\[
  \left.
  +\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],
  -\alpha^2 (S[c]S[c]^T+\alpha^2I)^{-1}r[c]\rangle\right.
\]
\[
  + \left.\alpha^2\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle\right)
\]
\[
  =  \frac{\alpha^2}{2}
\langle (S[c]S[c]^T+\alpha^2I)^{-1}r[c],r[c]\rangle 
\]
That is,
\begin{equation}
  \label{eqn:defwrialt}
  J_{\rm WRI}[c,d,q] = \frac{1}{2}\langle r[c], W_{\alpha}[c] r[c]\rangle
\end{equation}
with
\begin{equation}
  \label{eqn:defwriwt}
  W_{\alpha}[c] = \frac{\alpha^2}{2}(S[c]S[c]^T+\alpha^2I)^{-1}
\end{equation}
This remarkable identity shows that the WRI objective function is a {\em weighted norm of the data residual $r[c]$}.

@@

Define
\begin{equation}
  \label{eqn:defbeta}
\beta = \frac{z_{\rm max}-z_{\rm min}}{c_*^2} - 4 \alpha^2
\end{equation}
\begin{theorem}
  \label{thm:thm2}
  For $L$ as defined in \ref{eqn:defL}, and $\lambda <
  \lambda_0$, the minimizer of $J_{\rm WRI}[c,\dl,\wl \delta(\cdot-z_s)]$ on the
  complement of $[c_*-L\lambda, c_*+L\lambda]$ is
  \begin{itemize}
  \item $c=c_{\rm max}$ if $\beta<0$;
  \item $c=c_{\rm min}$ if $\beta>0$;
  \item any $c < c_*-L\lambda$ or $>c_*+L\lambda$ if $\beta=0$.
  \end{itemize}
\end{theorem}
\begin{proof}
  From \ref{eqn:wriscrewed}, \ref{eqn:wriwthom}, and
  \ref{eqn:fwiscrewed}, if $|c-c_*| > L\lambda$,
  \[
    J_{\rm WRI}[c,\dl,\wl\delta(\cdot-z_s)] =
    \frac{\alpha^2}{2}\left(\frac{z_{\rm max}-z_{\rm
          min}}{4c^2}+\alpha^2\right)^{-1}\frac{1}{2}\left(\frac{1}{4c^2}+\frac{1}{4c_*^2}\right)\|w_1\|^2
  \]
  \[
    =
    \frac{\alpha^2}{4}\frac{1+\frac{c^2}{c_*^2}}{z_{\rm 
        max}-z_{\rm min}+4c^2\alpha^2}
  \]
  This linear fractional function of $c^2$ is increasing, decreasing,
  or constant if $\beta>0, \beta<0$ or $\beta=0$, respectively.
\end{proof}

@@

\section{Discussion}
Theorems \ref{thm:thm1} and \ref{thm:thm2} call out the chief
conclusions of this work: that at least for the 1D acoustic
transmission inverse problem, both $J_{\rm FWI}[c,d,w]$ and
$J_{\rm WRI}[c,d,q]$ exhibit local minima (in $c$) far from the global
minimum for consistent data, and the domain of attraction of the
global minimizer can be arbitrarily small. These are striking
conclusions, but the 1D acoustic transmission problem is very special
and lacks fidelity to field practice. I shall show how these
approaches to solving this special problem share properties with a
much larger family of inversion methods. The theory developed to
explain these properties suggests methods that may not suffer the
non-convexity of FWI and WRI, and in fact do not in several cases that
I will mention.

First, a consequence of the results proven here: $J_{\rm FWI}[c,d,w]$
and $J_{\rm WRI}[c,d,q]$ are not smooth as joint functions of model
($c$) and data ($d$) vectors. If they were, their derivatives would be
bounded uniformly over bounded sets in $c,d$, but the two main results
show that this is not the case. As $\lambda \rightarrow 0$, $d=\dl$
varies within a ball $B \subset L^2([0,T])$ of radius $\|d_1\|$ (since
the $\|\dl\|$ is independent of $\lambda$), but the value of either
objective changes from a positive value (bounded away from zero
independently of $\lambda$) to zero over an interval of $c$ of length
$O(\lambda)$. Therefore the derivatives with respect to $c$ of both
objective functions are not bounded over
$[c_{\rm min},c_{\rm max}] \times B$.

While lack of smoothness is not in itself the most important property
established in the preceding sections, it is a necessary condition for
stable and reliable parameter recovery via local
optimization. Moreover, necessary conditions for smoothness are
known for a much wider class of quadratic form objectives for inverse
problems.
